Evolution of ORM Systems: Formal Foundations

### 2013-01-08 doc/description.tex
database transformation moved into section about transformations
appendValues renamed to insertValues

\subsection{Transformations for Manipulating a Database}
The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily in this paper.

\begin{align*}
	\mathbf{addColumn} : Label \times Column* \times Database \rightarrow Database \\ 	
	addColum(l, c, d) = \begin{cases}
		\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}\\ \\
		Inconsistent \;d 
	 \end{cases}\\ 
	 l \in Label, c = Column*, d = Database
\end{align*}

\begin{align*}
	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c }
	\end{cases} \\
	l_s,l_t,l_c \in Label, d = Database
\end{align*}

\begin{align*}
	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
	addTable(t, d) = \begin{cases}
		\inference{|t| > 1}{addTable(head(t),d) \bullet addTable(tail(t), d)}
\\ \\	
		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e)}{tables(d) \cup c}
\\ \\
		Inconsistent \;d 
	 \end{cases}\\ 
	t = Table*,  d = Database
\end{align*}

\begin{align*}
	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
	insertRow(r, d) = \begin{cases}
			\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r} \\ \\
			 Inconsistent \; d
		 \end{cases}	\\
		 r = Row*, d = Database
\end{align*}

\begin{align*}
	\mathbf{appendValues} : Label \times Pair* \times Database \rightarrow Daatabase \\
	appendValues(l, p, d) = \begin{cases}
 		\inference{|p| > 1}{appendValues(l, head(p), d)\bullet appendValues(l, tail(p), d)}  \\ \\
 		\inference{|p| = 1 \wedge \forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
	 	 Inconsistent \; d
 		\end{cases} \\
l \in Label, p = Pair*, d = Databases
\end{align*}

\begin{align*}
	\mathbf{copyValue} : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc} }{\begin{gathered}
			\forall p \in f_m(t_s, t_t)  : \\ appendValues(name(second(p)), \\ [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
		\end{gathered}}
 	\end{cases}
\end{align*}

@@

>>>>

@@

\subsection{Database Transformation}
The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily in this paper.
\paragraph{Add Column}
\begin{align*}
	\mathbf{addColumn} : Label \times Column* \times Database \rightarrow Database \\ 	
	addColum(l, c, d) = \begin{cases}
		\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}\\ \\
		Inconsistent \;d 
	 \end{cases}\\ 
	 l \in Label, c = Column*, d = Database
\end{align*}
\paragraph{COPY COLUMN}
\begin{align*}
	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c }
	\end{cases} \\
	l_s,l_t,l_c \in Label, d = Database
\end{align*}
\paragraph{Add Table}
\begin{align*}
	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
	addTable(t, d) = \begin{cases}
		\inference{|t| > 1}{addTable(head(t),d) \bullet addTable(tail(t), d)}
\\ \\	
		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e)}{tables(d) \cup c}
\\ \\
		Inconsistent \;d 
	 \end{cases}\\ 
	t = Table*,  d = Database
\end{align*}
\paragraph{Insert row}
\begin{align*}
	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
	insertRow(r, d) = \begin{cases}
			\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r} \\ \\
			 Inconsistent \; d
		 \end{cases}	\\
		 r = Row*, d = Database
\end{align*}
\paragraph{Insert Values}
\begin{align*}
	\mathbf{insertValues} : Label \times Pair* \times Database \rightarrow Daatabase \\
	insertValues(l, p, d) = \begin{cases}
 		\inference{|p| > 1}{insertValues(l, head(p), d)\bullet insertValues(l, tail(p), d)}  \\ \\
 		\inference{|p| = 1 \wedge \forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
	 	 Inconsistent \; d
 		\end{cases} \\
l \in Label, p = Pair*, d = Databases
\end{align*}
\paragraph{COPY VALUES}
\begin{align*}
	\mathbf{copyValue} : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc} }{\begin{gathered}
			\forall p \in f_m(t_s, t_t)  : \\ insertValues(name(second(p)), \\ [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
		\end{gathered}}
 	\end{cases}
\end{align*}
